---
title: "Untitled"
author: "Swapna Mahurkar-Joshi"
date: "October 26, 2017"
output: html_document
---

Extra from WGCNA

```{r}
#####################################################################################



# cluser samples check for outliers

```{r Chunk 3}
sampleTree = hclust(dist(datExpr0), method = "average");

# sizeGrWindow(12,9)
png("C:/Users/swapnajoshi-admin/Box Sync/DNA_methylation_R21_Analysis/temp/wgcna/plots/clusetringToDetectOutliers.png", width = 2000, height =1800, res = 200)
# par(cex = 0.6);
# par(mar = c(0,4,2,0))
plot(sampleTree, cex.axis = 1.5, cex.main = 2,
main = "Sample clustering to detect outliers",
cex.lab = 1.)
dev.off()

# Leave all the samples in
labels = cutreeStatic(sampleTree, cutHeight = 18)
keep = (labels==1)
datExpr <- datExpr0[keep, ]; dim(datExpr)

sampleTree = hclust(dist(datExpr), method = "average");
png("C:/Users/swapnajoshi-admin/Box Sync/DNA_methylation_R21_Analysis/temp/wgcna/plots/sampleClusteringNoOutlier.png", width = 2000, height = 1800, res = 200)
# par(cex = 0.6);
# par(mar = c(0,4,2,0))
plot(sampleTree, cex.axis = 1.5, cex.main = 2,
main = "Sample clustering to detect outliers",
cex.lab = 1.)
dev.off()

```

# Load clinical data

```{r Chunk 4, message=FALSE }
datTraits <- read.csv("C:/Users/swapnajoshi-admin/Box Sync/DNA_methylation_R21_Analysis/temp/functional_norm/Metadata_colon_methylation.csv", row.names = 1); dim(datTraits)
# datTraits.IBS <- subset(datTraits, datTraits$Group==2); dim(datTraits.IBS)
datExpr1 <- datExpr[row.names(datExpr)%in%row.names(datTraits),]; dim(datExpr1)
datTraits1 <- datTraits[row.names(datTraits)%in%row.names(datExpr1),];
datTraits1 <- datTraits1[row.names(datExpr1),];dim(datTraits1)
datTraits <- datTraits1[,c(1:21,25,27)]; dim(datTraits)
datTraits <- as.data.frame(datTraits)
datExpr <-datExpr1
match(row.names(datTraits), row.names(datExpr1))
collectGarbage()

datTraits$BH_Colon_Exam_num <-  ifelse(datTraits$BH_Colon_Exam=="C",1,ifelse(datTraits$BH_Colon_Exam=="D",2,ifelse(datTraits$BH_Colon_Exam=="M",6,ifelse(datTraits$BH_Colon_Exam=="U",6,4))))
datTraits$BH_Colon_Exam <- NULL

datTraits1 <- datTraits
datTraits[,23] <- as.numeric(as.character(datTraits[,23]))
datTraits <- apply(datTraits,2,as.numeric)
row.names(datTraits) <- row.names(datTraits1)
colnames(datTraits) <- colnames(datTraits1)
datTraits <- as.data.frame(datTraits)
save(datTraits, file = "C:/Users/swapnajoshi-admin/Box Sync/DNA_methylation_R21_Analysis/temp/wgcna/datTraits.col.rda")

match(row.names(datTraits),row.names(datExpr))
collectGarbage()
rm(list = c("datExpr0","datTraits1", "datExpr1"))
```

```{r}
sampleTree2 = hclust(dist(datExpr), method = "average")
traitColors = numbers2colors(datTraits, signed = FALSE,
naColor="grey")
datTraits <- as.data.frame(datTraits)
lab1 <- ifelse(datTraits$Group==1,"HC", "IBS")
png("C:/Users/swapnajoshi-admin/Box Sync/DNA_methylation_R21_Analysis/temp/wgcna/plots/heirarchical_clustering_col.png", height = 3000, width = 3200, res=200)
plotDendroAndColors(sampleTree2, traitColors,groupLabels = names(datTraits),dendroLabels = lab1, cex.dendroLabels = 0.8, main = "Sample dendrogram, trait heatmap")
dev.off()
save(datExpr, datTraits, file = "C:/Users/swapnajoshi-admin/Box Sync/DNA_methylation_R21_Analysis/temp/wgcna/IBS-01-dataInput-colon.RData")

```
# Choosing soft thresholding power β
```{r}
disableWGCNAThreads()
powers = c(c(1:10), seq(from = 12, to=20, by=2))
sft = pickSoftThreshold(datExpr, powerVector = powers,
verbose = 5)

save(sft , file = "C:/Users/swapnajoshi-admin/Box Sync/DNA_methylation_R21_Analysis/temp/wgcna/sft_power12.rda")
```


```{r}
# sizeGrWindow(9, 5)
png("C:/Users/swapnajoshi-admin/Box Sync/DNA_methylation_R21_Analysis/temp/wgcna/plots/soft_power_col.png", height = 3000, width = 3200, res=200)
par(mfrow = c(1,2))
cex1 = 0.9
plot(sft$fitIndices[,1],
-sign(sft$fitIndices[,3])*sft$fitIndices[,2],
xlab="Soft Threshold (power)",
ylab="Scale Free Topology Model Fit,signed R^2",type="n",
main = paste("Scale independence"));
text(sft$fitIndices[,1],
-sign(sft$fitIndices[,3])*sft$fitIndices[,2],
labels=powers,cex=cex1,col="red");
# this line corresponds to using an R^2 cut-off of h
abline(h=0.872,col="red")
dev.off()
```

```{r}
# sizeGrWindow(9, 5)
png("C:/Users/swapnajoshi-admin/Box Sync/DNA_methylation_R21_Analysis/temp/wgcna/plots/power_connectivity_colon.png", height = 3000, width = 3200, res=200)
par(mfrow = c(1,2))
cex1 = 0.9;
plot(sft$fitIndices[,1], sft$fitIndices[,5],
xlab="Soft Threshold (power)",
ylab="Mean Connectivity", type="n",
main = paste("Mean connectivity"))
text(sft$fitIndices[,1], sft$fitIndices[,5],
labels=powers, cex=cex1,col="red")
dev.off()
```

```{r}
net = blockwiseModules(datExpr, power = 4,
TOMType = "unsigned",
minModuleSize = 30,
reassignThreshold = 0,
mergeCutHeight = 0.25,
numericLabels = TRUE,
pamRespectsDendro = FALSE,
saveTOMs = TRUE,
saveTOMFileBase="IBS_TOM",
verbose = 3)
```

```{r}
png("C:/Users/swapnajoshi-admin/Box Sync/DNA_methylation_R21_Analysis/temp/wgcna/plots/heirarchical_colors.png", height = 3000, width = 3200, res=200)
plotDendroAndColors(net$dendrograms[[1]],
net$blockGenes[[1]],
"Module colors",
dendroLabels = FALSE, hang = 0.03,
addGuide = TRUE, guideHang = 0.05)
dev.off()
```


```{r}
moduleLabels = net$colors
moduleColors = labels2colors(net$colors)
MEs = net$MEs;
geneTree = net$dendrograms[[1]];
save(MEs, moduleLabels, moduleColors, geneTree,
file="C:/Users/swapnajoshi-admin/Box Sync/DNA_methylation_R21_Analysis/temp/wgcna/IBS-02-networkConstruction-auto-col.RData")
```

```{r}
# Define numbers of genes and samples
nGenes = ncol(datExpr);
nSamples = nrow(datExpr);
# Recalculate MEs with color labels
MEs0 = moduleEigengenes(datExpr, moduleColors)$eigengenes
MEs = orderMEs(MEs0)
moduleTraitCor = cor(MEs, datTraits, use = "p");
moduleTraitPvalue = corPvalueStudent(moduleTraitCor, nSamples)
```

```{r}
datTraits <- as.data.frame(datTraits)
textMatrix =  paste(signif(moduleTraitCor, 2), "\n(",signif(moduleTraitPvalue,
1), ")", sep = "");
dim(textMatrix) = dim(moduleTraitCor)
png("C:/Users/swapnajoshi-admin/Box Sync/DNA_methylation_R21_Analysis/temp/wgcna/plots/moduleTraitRelationship_col.png", height = 2200, width = 2500, res= 200)
par(mar = c(6, 8.5, 3, 3));
labeledHeatmap(Matrix = moduleTraitCor,
xLabels = names(datTraits),
yLabels = names(MEs),
ySymbols = names(MEs),
colorLabels = FALSE,
colors = blueWhiteRed(50),
textMatrix = textMatrix,
setStdMargins = FALSE,
cex.text = 0.6,
zlim = c(-1,1),
main = paste("Module-trait relationships in colon"))
dev.off()
```
# Identify important modules
```{r}
moduleTraitCor1 <- as.data.frame(moduleTraitCor)
m1 <- moduleTraitCor1[apply(moduleTraitCor1, MARGIN = 1, function(x) any(x < -0.16 | x>0.16)), ]
m2 <- m1[ ,apply(m1, MARGIN = 2, function(x) any(x < -0.16 | x>0.16))]

moduleTraitPvalue1 <- as.data.frame(moduleTraitPvalue)
p1 <- moduleTraitPvalue1[apply(moduleTraitPvalue1, MARGIN = 1, function(x) any(x < 0.05)), ]
p2 <- p1[ ,apply(p1, MARGIN = 2, function(x) any(x < 0.05))]

m2 <- moduleTraitCor1[row.names(moduleTraitCor1)%in%row.names(p2), colnames(moduleTraitCor1)%in%colnames(p2)]

textMatrix1 =  paste(m2, "\n(",p2,")", sep = "");

png("C:/Users/swapnajoshi-admin/Box Sync/DNA_methylation_R21_Analysis/temp/wgcna/plots/moduleTraitRelationship_col_sigOnly.png", height = 2200, width = 2500, res= 200)
par(mar = c(6, 8.5, 3, 3));
labeledHeatmap(Matrix = m2,
xLabels = colnames(m2),
yLabels = row.names(m2),
ySymbols = row.names(m2),
colorLabels = FALSE,
colors = blueWhiteRed(50),
textMatrix = textMatrix1,
setStdMargins = FALSE,
cex.text = 0.9,
zlim = c(-1,1),
main = paste("Significant Module-trait relationships in colon"))
dev.off()

# Yellow = BSQ_AbdPain
# black = BSQ_OverallSx
# turqoise = ACE
# brown = ACE

```

```{r}
# Module of interest:  yellow; trait = BSQ_AbdPain
# Define variable abd pain containing column of datTrait
pain = as.data.frame(datTraits$BSQ_AbdPain);
names(pain) = "AbdominalPain"
modNames = substring(names(MEs), 3)
geneModuleMembership = as.data.frame(cor(datExpr, MEs, use = "p"));
MMPvalue = as.data.frame(corPvalueStudent(as.matrix(geneModuleMembership), nSamples));
names(geneModuleMembership) = paste("MM", modNames, sep="");
names(MMPvalue) = paste("p.MM", modNames, sep="");
geneTraitSignificance = as.data.frame(cor(datExpr, pain, use = "p"));
GSPvalue = as.data.frame(corPvalueStudent(as.matrix(geneTraitSignificance), nSamples));
names(geneTraitSignificance) = paste("GS.", names(pain), sep="");
names(GSPvalue) = paste("p.GS.", names(pain), sep="");


# Identify genes with high GS and MM
module = "yellow"
column = match(module, modNames);
moduleGenes = moduleColors==module;
# sizeGrWindow(7, 7);
png("wgcna/plots/AbdominalPain_yellow_colon.png", height = 2000, width = 2500, res= 200)
par(mfrow = c(1,1));
verboseScatterplot(
abs(geneModuleMembership[moduleGenes,
column]),
abs(geneTraitSignificance[moduleGenes, 1]),
xlab = paste("Module Membership in",module,"module"),
ylab = "Gene significance for abdominal pain",
main = paste("Module membership vs. gene significance\n"),
cex.main = 1.2, cex.lab = 1.2,
cex.axis = 1.2, col = module)
dev.off()

# cpg sites in yellow module
load("hm450annotations.rda")

hm450annotationsYellow <- hm450annotations[row.names(hm450annotations)%in%row.names(MMPvalue[moduleGenes,]),]
write.csv(hm450annotationsYellow, file = "wgcna/abdominalPainYellowAnnotaionsColon.csv")

hm450annotations <- as.data.frame(hm450annotations)
geneInfo0 = data.frame(Probes = colnames(datExpr),
geneSymbol =hm450annotations[row.names(hm450annotations)%in%colnames(datExpr), 37],
RelationToIsland = hm450annotations[row.names(hm450annotations)%in%colnames(datExpr), 19],
moduleColor = moduleColors,
geneTraitSignificance,
GSPvalue)

# Order modules by their significance for weight
modOrder = order(-abs(cor(MEs, pain, use = "p")));
# Add module membership information in the chosen order
for (mod in 1:ncol(geneModuleMembership))
{
oldNames = names(geneInfo0)
geneInfo0 = data.frame(geneInfo0, geneModuleMembership[, modOrder[mod]],
MMPvalue[, modOrder[mod]]);
names(geneInfo0) = c(oldNames, paste("MM.", modNames[modOrder[mod]], sep=""),
paste("p.MM.", modNames[modOrder[mod]], sep=""))
}
# Order the genes in the geneInfo variable first by module color, then by geneTraitSignificance
geneOrder = order(geneInfo0$moduleColor, -abs(geneInfo0$GS.AbdominalPain));
geneInfo = geneInfo0[geneOrder, ]

hm450annotationsyellow <- hm450annotations[row.names(hm450annotations)%in%names(datExpr)[moduleColors=="yellow"],]
write.csv(hm450annotationsYellow, file = "wgcna/abdominalPainYellowAnnotaionsColon.csv")
write.csv(geneInfo, file = "wgcna/geneInfo_abdominalPain.csv")
fix(geneInfo)

```

```{r}
# Module of interest:  black; trait = BSQ_OverallSx
# Define variable  containing column of datTrait
severity = as.data.frame(datTraits$BSQ_OverallSx);
names(severity) = "OverallSeverity"
modNames = substring(names(MEs), 3)
geneModuleMembership = as.data.frame(cor(datExpr, MEs, use = "p"));
MMPvalue = as.data.frame(corPvalueStudent(as.matrix(geneModuleMembership), nSamples));
names(geneModuleMembership) = paste("MM", modNames, sep="");
names(MMPvalue) = paste("p.MM", modNames, sep="");
geneTraitSignificance = as.data.frame(cor(datExpr,  severity, use = "p"));
GSPvalue = as.data.frame(corPvalueStudent(as.matrix(geneTraitSignificance), nSamples));
names(geneTraitSignificance) = paste("GS.", names(severity), sep="");
names(GSPvalue) = paste("p.GS.", names(severity), sep="");

# Identify genes with high GS and MM

module = "black"
column = match(module, modNames);
moduleGenes = moduleColors==module;
# sizeGrWindow(7, 7);
png("wgcna/plots/overallSeverity_black_colon.png", height = 2000, width = 2500, res= 200)
par(mfrow = c(1,1));
verboseScatterplot(
abs(geneModuleMembership[moduleGenes,
column]),
abs(geneTraitSignificance[moduleGenes, 1]),
xlab = paste("Module Membership in",module,"module"),
ylab = "Gene significance for OverallSeverity",
main = paste("Module membership vs. gene significance\n"),
cex.main = 1.2, cex.lab = 1.2,
cex.axis = 1.2, col = module)
dev.off()

# cpg sites in black module

datExpr <- as.data.frame(datExpr)
hm450annotationsblack <- hm450annotations[row.names(hm450annotations)%in%names(datExpr)[moduleColors=="black"],]

hm450annotations <- as.data.frame(hm450annotations)
geneInfo0 = data.frame(Probes = colnames(datExpr),
geneSymbol =hm450annotations[row.names(hm450annotations)%in%colnames(datExpr), 37],
RelationToIsland = hm450annotations[row.names(hm450annotations)%in%colnames(datExpr), 19],
moduleColor = moduleColors,
geneTraitSignificance,
GSPvalue)

# Order modules by their significance for weight
modOrder = order(-abs(cor(MEs, pain, use = "p")));
# Add module membership information in the chosen order
for (mod in 1:ncol(geneModuleMembership))
{
oldNames = names(geneInfo0)
geneInfo0 = data.frame(geneInfo0, geneModuleMembership[, modOrder[mod]],
MMPvalue[, modOrder[mod]]);
names(geneInfo0) = c(oldNames, paste("MM.", modNames[modOrder[mod]], sep=""),
paste("p.MM.", modNames[modOrder[mod]], sep=""))
}
# Order the genes in the geneInfo variable first by module color, then by geneTraitSignificance
geneOrder = order(geneInfo0$moduleColor, -abs(geneInfo0$GS.OverallSeverity));
geneInfo = geneInfo0[geneOrder, ]


hm450annotationsblack <- hm450annotations[row.names(hm450annotations)%in%names(datExpr)[moduleColors=="black"],]

write.csv(hm450annotationsblack, file = "wgcna/OverallSevBlackAnnotaionsColon.csv")
write.csv(geneInfo , file = "wgcna/geneInfo_overallSeverity_black.csv")
fix(geneInfo)


```


```{r}
# Module of interest:  turqoise; trait = ACE
# Define variable  containing column of datTrait
ACE = as.data.frame(datTraits$ACE_Score);
names(ACE) = "ACE"
modNames = substring(names(MEs), 3)
geneModuleMembership = as.data.frame(cor(datExpr, MEs, use = "p"));
MMPvalue = as.data.frame(corPvalueStudent(as.matrix(geneModuleMembership), nSamples));
names(geneModuleMembership) = paste("MM", modNames, sep="");
names(MMPvalue) = paste("p.MM", modNames, sep="");
geneTraitSignificance = as.data.frame(cor(datExpr,  ACE, use = "p"));
GSPvalue = as.data.frame(corPvalueStudent(as.matrix(geneTraitSignificance), nSamples));
names(geneTraitSignificance) = paste("GS.", names(ACE), sep="");
names(GSPvalue) = paste("p.GS.", names(ACE), sep="");

# Identify genes with high GS and MM

module = "turquoise"
column = match(module, modNames);
moduleGenes = moduleColors==module;
# sizeGrWindow(7, 7);
png("wgcna/plots/ACE_turquoise_colon.png", height = 2000, width = 2500, res= 200)
# par(mfrow = c(1,1));
verboseScatterplot(
abs(geneModuleMembership[moduleGenes,
column]),
abs(geneTraitSignificance[moduleGenes, 1]),
xlab = paste("Module Membership in",module,"module"),
ylab = "Gene significance for ACE",
main = paste("Module membership vs. gene significance\n"),
cex.main = 1.2, cex.lab = 1.2,
cex.axis = 1.2, col = module);
dev.off()

# cpg sites in black module

datExpr <- as.data.frame(datExpr)
hm450annotationsblack <- hm450annotations[row.names(hm450annotations)%in%names(datExpr)[moduleColors=="turquoise"],]

hm450annotations <- as.data.frame(hm450annotations)
geneInfo0 = data.frame(Probes = colnames(datExpr),
geneSymbol =hm450annotations[row.names(hm450annotations)%in%colnames(datExpr), 37],
RelationToIsland = hm450annotations[row.names(hm450annotations)%in%colnames(datExpr), 19],
moduleColor = moduleColors,
geneTraitSignificance,
GSPvalue)

# Order modules by their significance for weight
modOrder = order(-abs(cor(MEs, pain, use = "p")));
# Add module membership information in the chosen order
for (mod in 1:ncol(geneModuleMembership))
{
oldNames = names(geneInfo0)
geneInfo0 = data.frame(geneInfo0, geneModuleMembership[, modOrder[mod]],
MMPvalue[, modOrder[mod]]);
names(geneInfo0) = c(oldNames, paste("MM.", modNames[modOrder[mod]], sep=""),
paste("p.MM.", modNames[modOrder[mod]], sep=""))
}
# Order the genes in the geneInfo variable first by module color, then by geneTraitSignificance
geneOrder = order(geneInfo0$moduleColor, -abs(geneInfo0$GS.ACE));
geneInfo = geneInfo0[geneOrder, ]


hm450annotationsturquoise <- hm450annotations[row.names(hm450annotations)%in%names(datExpr)[moduleColors=="turquoise"],]

write.csv(hm450annotationsturquoise, file = "wgcna/ACEturquoiseAnnotaionsColon.csv")
write.csv(geneInfo , file = "wgcna/geneInfo_ACE_turquoise.csv")
fix(geneInfo)


```
##############################################################################################
##############################################################################################
##############################################################################################

# Install Mergeomics
```{r}
source("https://bioconductor.org/biocLite.R")
biocLite("Mergeomics")
```

# Import library scripts.
```{r}
library(Mergeomics)
```
###### MSEA (Marker set enrichment analysis)  ###

```{r}
job.msea <- list()
job.msea$label <- "hdlc"
job.msea$folder <- "Results"
job.msea$genfile <- system.file("extdata", 
"genes.hdlc_040kb_ld70.human_eliminated.txt", package="Mergeomics")
job.msea$marfile <- system.file("extdata", 
"marker.hdlc_040kb_ld70.human_eliminated.txt", package="Mergeomics")
job.msea$modfile <- system.file("extdata", 
"modules.mousecoexpr.liver.human.txt", package="Mergeomics")
job.msea$inffile <- system.file("extdata", "coexpr.info.txt", 
package="Mergeomics")
job.msea$nperm <- 100 ## default value is 20000 (this is recommended)
job.msea <- ssea.start(job.msea)
job.msea <- ssea.prepare(job.msea)
job.msea <- ssea.control(job.msea)
job.msea <- ssea.analyze(job.msea)
job.msea <- ssea.finish(job.msea)
```
######  Create intermediary datasets for KDA ###########
```{r}
syms <- tool.read(system.file("extdata", "symbols.txt", 
package="Mergeomics"))
syms <- syms[,c("HUMAN", "MOUSE")]
names(syms) <- c("FROM", "TO")

## default and recommended rmax=0.33.
## min.module.count is the number of the pathways to be taken from the MSEA
## results to merge. If it is not specified (NULL), all the pathways having 
## MSEA-FDR value less than 0.25 will be considered for merging if they are 
## overlapping with the given ratio rmax. 
job.kda <- ssea2kda(job.msea, rmax=0.2, symbols=syms, min.module.count=NULL) 
```
######   wKDA (Weighted key driver analysis)    ##########
```{r}
job.kda$netfile <- system.file("extdata", 
"network.mouseliver.mouse.txt", package="Mergeomics")
job.kda <- kda.configure(job.kda)
job.kda <- kda.start(job.kda)
job.kda <- kda.prepare(job.kda)
job.kda <- kda.analyze(job.kda)
job.kda <- kda.finish(job.kda)
```
######  Prepare network files for visualization   #########
```{r}
## Creates the input files for Cytoscape (http://www.cytoscape.org/)
job.kda <- kda2cytoscape(job.kda)
```



```

